---
title: "PA1_template"
author: "Jose Alberto"
date: "22 de octubre de 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##Code for reading data
```{r fig.height=4.5,fig.width=4.5, fig.align='center'}
datos<-read.csv("C:\\Users\\juanc\\Desktop\\activity.csv")
```

##Transforming data type of dates as Date
```{r fig.height=4.5,fig.width=4.5, fig.align='center'}
fecha<-as.character(datos$date)
datos$date<-as.Date(fecha)
```

##Sum of steps per day 
```{r fig.height=4.5,fig.width=4.5, fig.align='center'}
steps<- aggregate(datos$steps ~ datos$date, datos, sum)
```

##Change column names
```{r fig.height=4.5,fig.width=4.5, fig.align='center'}
names(steps) <- c("fecha", "pasos")
```

##Histogram of steps per day (on average)
```{r fig.height=4.5,fig.width=4.5, fig.align='center'}
hist(steps$pasos, main = paste("Total Steps Each Day"), col="purple", xlab="Number of Steps")
```

##Mean and Median of steps per day
```{r fig.height=4.5,fig.width=4.5, fig.align='center'}
rmean <- mean(steps$pasos) 
rmedian <- median(steps$pasos)
```

##Mean of steps per interval
```{r fig.height=4.5,fig.width=4.5, fig.align='center'}
steps2<- aggregate(datos$steps ~ datos$interval, datos,mean)
names(steps2) <- c("intervalo", "pasos")
```

##Time series of number of steps on average per interval
```{r fig.height=4.5,fig.width=4.5, fig.align='center'}
plot(steps2$intervalo,steps2$pasos,main="Average number of steps each interval", type="l", col="blue")
```

##Interval which contains the max number of steps on average per interval
```{r fig.height=4.5,fig.width=4.5, fig.align='center'}
steps2[which.max(steps2$pasos),1]
max(steps2$pasos)
maximo <- subset(steps2, pasos>=max(steps2$pasos), select = intervalo)
maximo
```

##Counting NA values in steps field
```{r fig.height=4.5,fig.width=4.5, fig.align='center'}
nulos <- is.na(datos$steps)
sum(nulos)
```
##We create a new data frame which will not contains NA values in steps field
```{r fig.height=4.5,fig.width=4.5, fig.align='center'}
copia <- datos
```
##Replacing NA values by mean of steps in corresponding both interval and day
```{r fig.height=4.5,fig.width=4.5, fig.align='center'}
copia$steps<- ifelse(is.na(copia$steps),steps2$pasos[match(copia$interval,steps2$intervalo)],copia$steps)
```
##Verify that our new data frame doesn´t contain NA values
```{r fig.height=4.5,fig.width=4.5, fig.align='center'}
nulos2 <- is.na(copia$steps)
sum(nulos2)
```
##Transforming data type of dates as Date in new data frame
```{r fig.height=4.5,fig.width=4.5, fig.align='center'}
fecha2<-as.character(copia$date)
copia$date<-as.Date(fecha2)
steps3<- aggregate(copia$steps ~ copia$date, copia, sum)
names(steps3) <- c("fecha", "pasos")
```
##Histogram of steps per day (on average) in new data frame
```{r fig.height=4.5,fig.width=4.5, fig.align='center'}
hist(steps3$pasos, main = paste("Total Steps Each Day"), col="green", xlab="Number of Steps")
```
##Finding day name for each date
```{r fig.height=4.5,fig.width=4.5, fig.align='center'}
fechas<- as.Date(copia$date)
dias<-weekdays(fechas)
dias<-ifelse(grepl("Sabado", dias), "Fin de Semana", dias)
dias<-ifelse(grepl("Domingo", dias), "Fin de Semana", dias)
dias<-ifelse(grepl("Lunes", dias), "Entre semana", dias)
dias<-ifelse(grepl("Martes", dias), "Entre semana", dias)
dias<-ifelse(grepl("Miercoles", dias), "Entre semana", dias)
dias<-ifelse(grepl("Jueves", dias), "Entre semana", dias)
dias<-ifelse(grepl("Viernes", dias), "Entre semana", dias)
dias<- as.factor(dias)
```

##Adding day names in data frame which doesn't contain NA values
```{r fig.height=4.5,fig.width=4.5, fig.align='center'}
copia<-cbind(copia,dias)
```

##Importing libraries needed to plot following graphs
```{r}
library(magrittr)
library(dplyr)
library(ggplot2)
```

##New data frame with day names
```{r fig.height=4.5,fig.width=4.5, fig.align='center'}
porTipoDeDia <- copia %>% group_by(interval, dias) %>% summarise(steps = mean(steps))
```


##Plotting graphs of number of steps (on average) for each day of the week
```{r, fig.align='center'}
dev.off()
ggplot(porTipoDeDia, aes(x=interval, y=steps, color = dias)) + geom_line() + facet_wrap(~dias, nrow=7, ncol=1)
```
